Linear regression for approximating real and synthetic data

### 2022-10-10 doc/paper.tex
fix introduction

% One of the core challenges of machine learning is model selection, because we cannot make a priori assessments about how well the model will fit our dataset. The model is at the heart of machine learning, its like the CPU of the "machine" in machine learning. It really puts the "machine" in machine learning, so to speak. Learning is like the RAM. We are the keyboard and you are the screen.
Machine learning is well suited for finding complex relationships that are impossible to describe as a series of logical statements, or program by hand. Machine learning methods can look at millions of datapoints and parameters at once, and thus find links in the data which could never be found by manual inspection. But to do this, we require a model that is suited to our dataset. However, in order to create an optimal model, we require optimal parameters which cannot be known a priori. Thus, we require extensive testing and tuning of parameters to inform our selection. In this paper, we explore

% One of the core challenges of machine learning is model selection, because we cannot make a priori assessments about how well the model will fit our dataset. If our model is not complex enough for our problem, we are unlikely to reach a good prediction. Likewise, if our model is too complex, we run the risk of overfitting, where our model adapts to our training data perfectly and thus loses the ability to make predictions on unseen data. Given the enormous datasets that

% Linear regression is a simple method in machine learning, used to create a fit of a dataset of which it can produce continuous predictions. This is achieved by assuming that we can model the output values as a linear combination of continuous functions of the input values. For example, we can assume that the observed values in our dataset can be approximated by different polynomial degrees of the input values and find the linear coefficients $\beta$ which minimize the difference between our observed and predicted values. This general idea that we can create a model which approximates our observed data with a minimal difference between observed and predicted values is a problem at the very heart of machine learning.

% Furthermore, we will compare different models of linear regression, namely Ordinary Least Squares (OLS), Ridge and Lasso regression, and employ resampling methods, namely bootstrap and cross-validation to ensure that we draw conclusive results as these help us reduce the randomness in our parameters based off how we split our data. <source?>

>>>>

One of the core challenges in machine learning is model selection, because we cannot make a priori assessments about how well the model will fit our dataset. A multitude of factors determine the efficacy of our model, the machine learning method we use, the hyperparameters of said method, and the model complexity we select. An uninformed decision in model selection could result in an under- or overfitting to the data, yielding poor results. Thus, it is important that we rigorously test our tuning such that these factors coincide at an intersection, providing the minimal error in unity.

In this paper, we will provide a detailed inspection of the Ordinary Least Squares (OLS), Ridge and Lasso linear regression methods, it's hyperparameters and varying model complexity with the aim of making an informed decision about model selection when fitting to different datasets. The datasets consist of synthetic data created by the Franke function, and real datasets of terrain in Norway. We inform our decisions by plotting the mean squared error (MSE), our cost function, over different polynomial degrees describing model complexity, over different resampling techniques and over different values of lambda, which is the hyperparameter responsible for a suitable regularization. We will perform a bias-variance decomposition for the different linear regression methods to check for under- and overfitting and plot our implementations of the methods up against popular implementations such as scikit learn to ensure that the behaviour of our models is expected.

First, we will go more into depth about the methods and the theory behind them, along with a brief description of our implementation. We will further continue to the results, followed up by a critical discussion of the results. Finally, we will give a conclusion on the optimal model selection for the different datasets.
